---
title: "Relative testes size"
author: "Annemarie van der Marel"
date: "2022-12-15"
output: html_document
---


```{r}
sessionInfo()
```

remove marine mammals (cetacea & sirenia) and humans, as they are outliers


# Load libraries
```{r setup, include=FALSE}
# data cleaning
library(tidyverse) 

#plotting
library(viridis)
library(RColorBrewer)

```


# Import  data
```{r}

# data with Baker 2019 dataset and crossreffed with vertlife
full_df <- read.csv("../data/testes_lifehistory_mating_care_Dec17.csv") %>%
  arrange(Binomial)

length(unique(full_df$Binomial)) 
multiple_entries <- full_df %>% 
  group_by(Binomial) %>%
  tally() %>%
  filter(n>1) # 43 species with > 1 entry

# species for phylogenetic tree
species631 <- unique(full_df$Binomial)
#write.csv(species631, "../data/631species.csv")

```
## remove outliers
remove marine mammals (cetacea & sirenia) and humans, as they are outliers
```{r}

unique(full_df$order)

df <- full_df %>%
  filter(!order %in% c("Cetacea", "Sirenia"), 
         Binomial!="Homo_sapiens")
  
unique(df$order)
```



## number of species per trait

dataset to use (excludes marine mammal orders such as Cetacea and Sirenia and humans):
In total, we have 571 terrestrial mammal species in 17 orders for which we can calculate relative testes size. Of these 571, we have 541 species with litter size data, 460 with litters per year, 469 with longevity, 373 with mating system, and 263 with paternal care data. We have 220 species where we have data for all our predictor variables.


```{r species with data}
all_df <- df

n_distinct(all_df$Binomial) # number of species with testes data
n_distinct(all_df$order)

n_ls <- all_df %>%
  filter(!is.na(litter_size)) 
n_distinct(n_ls$Binomial)

n_ly <- all_df %>%
  filter(!is.na(litter_year))
n_distinct(n_ly$Binomial)

n_lg <- all_df %>%
  filter(!is.na(longevity))
n_distinct(n_lg$Binomial)

n_ms <- all_df %>%
  filter(!is.na(mating_system))
n_distinct(n_ms$Binomial)

n_pc <- all_df %>%
  filter( paternal_care!="")
n_distinct(n_pc$Binomial)

n_all <- na.omit(all_df) 
n_distinct(n_all$Binomial)

```


full dataset includng Baker et al 2019 dataset:
In total, we have 631 species for which we can calculate relative testes size. Of these 631, we have 593 species with litter size data, 504 with litters per year, 517 with longevity, 381 with mating system, and 264 with paternal care data. We have 257 species where we have data for all our predictor variables.
```{r species with data}
all_df <- full_df

n_distinct(all_df$Binomial) # number of species with testes data

n_testes <- all_df %>%
  filter(is.na(testes_g)) 

n_ls <- all_df %>%
  filter(!is.na(litter_size)) 
n_distinct(n_ls$Binomial)

n_ly <- all_df %>%
  filter(!is.na(litter_year))
n_distinct(n_ly$Binomial)

n_lg <- all_df %>%
  filter(!is.na(longevity))
n_distinct(n_lg$Binomial)

n_ms <- all_df %>%
  filter(!is.na(mating_system))
n_distinct(n_ms$Binomial)

n_pc <- all_df %>%
  filter( paternal_care!="")
n_distinct(n_pc$Binomial)

n_all <- na.omit(all_df) 
n_distinct(n_all$Binomial)

```



Small dataset excludng Baker et al 2019 dataset:
In total, we have 210 species for which we can calculate relative testes size. Of these 210, we have 198 species with litter size data, 147 with litters per year, 170 with longevity, 108 with mating system, and 108 with paternal care data. We only have 8 species (all squirrels) where we have data for all our predictor variables. 

```{r missing species}
n_testes <- all_df %>%
  filter(is.na(testes_g)) 
n_distinct(all_df$Binomial)

n_ls <- all_df %>%
  filter(is.na(LS)) 
n_distinct(n_ls$Binomial)
length(unique(check_df$Binomial))

n_ly <- all_df %>%
  filter(is.na(LY))
n_distinct(n_ly$Binomial)

n_lg <- all_df %>%
  filter(is.na(LG))
n_distinct(n_lg$Binomial)

n_ms <- all_df %>%
  filter(is.na(mating_system))
n_distinct(n_ms$Binomial)

n_pc <- all_df %>%
  filter( paternal_care=="")
n_distinct(n_pc$Binomial)

n_all <- na.omit(all_df) 
n_distinct(n_all)
n_order <- n_all %>%
  group_by(order) %>%
  tally()


```

 

# plot testes ~ bm

```{r set colors by order}

n_distinct(all_df$order)

# 26 qualitatively different colors from Dave Keenan
colors26 <-  c("#00B7FF", "#004DFF", "#00FFFF", "#826400", "#580041", "#FF00FF", "#00FF00",
               "#C500FF", "#B4FFD7", "#FFCA00", "#969600", "#B4A2FF", "#C20078", "#0000C1",
               "#FF8B00", "#FFC8FF", "#666666", "#FF0000", "#CCCCCC", "#009E8F", "#D7A870", 
               "#8200FF", "#960000", "#BBFF00",  "#006F00", "#1f78b4", "#8E0152") # "#1f78b4", "#8E0152", "#FFFF00", are the extr 2 colors

set.seed(42)
colorXorder <-  sample(colors26, n_distinct(all_df$order))

```


```{r}
ggplot(all_df, aes(x=log(BM_g), y=log(testes_g))) +
  geom_point(aes(color=order)) +
  labs(x="log body mass (g)", y="log testes mass (g)") +
  theme_classic() +
  scale_colour_manual(values = colorXorder) 

ggsave("../figures/log_testes_bm.pdf") 

# without legend
ggplot(all_df, aes(x=log(BM_g), y=log(testes_g))) +
  geom_point(aes(color=order)) +
  labs(x="log body mass (g)", y="log testes mass (g)") +
  theme_classic() +
  theme(legend.position = "none") +
  scale_colour_manual(values = colorXorder) 

ggsave("../figures/log_testes_bm_nolegend.pdf") 
```

How many species by order?
```{r}

n_order <- all_df %>%
  group_by(order) %>%
  tally()
```



# Relative testes size
Kenagy & Trombulak 1986
Testes mass = 0.035 x body mass^0.72
The equation is Y=0.035X to the power of 0.72, Where the mass of the testes is Y.  The body mass is X.

The relative testes size is the ratio of observed testes size to the testes size predicted by this equation

from thesis Anjos: 
Richardson's ground squirrels = 0.61 ± 0.06
```{r rts function}

check_outliers <- all_df %>%
  summarize(checkmass= ifelse(testes_g==BM_g, "same", "no"))
# testes mass incorrect: should be 1.2 instead of 1021; already corrected
# all_df %>%   filter(Binomial=="Aotus_trivirgatus") 
# all_df$testes_g[all_df$Binomial=="Aotus_trivirgatus"] <- 1.2


rts_df <- all_df %>%
  mutate(exp.testes=0.035*(BM_g^0.72),
         rts=testes_g/exp.testes) %>%
  dplyr::select(X, order, family, common_name, Binomial, rts, 
                exp.testes, testes_g,log.testes., ref_testes, everything())


head(rts_df)

# include rts for which we only had relative testes size and not testes_mass
missingrts <- rts_df %>% 
  filter(is.na(rts))
# rts_df %>%
#   filter(Binomial=="Ammospermophilus_leucurus")
# rts_df$rts[rts_df$Binomial=="Ammospermophilus_leucurus"] <- 2.24

rts_df %>%
  filter(Binomial=="Urocitellus_richardsonii")
rts_df$rts[rts_df$Binomial=="Urocitellus_richardsonii"] <- 0.61

# rts_df <- rts_df %>%
#   dplyr::select(-rel_testes_size) 


#write.csv(rts_df, "../data/rts_life_mating_care.csv")
```


## Check for outliers in relative testes size data
```{r}

#rts_df <- read.csv("../data/rts_life_mating_care.csv")



check_outliers <- rts_df %>%
  summarize(checkmass= ifelse(testes_g>BM_g, "same", "no"))
unique(check_outliers$checkmass)

range(rts_df$rts) #order cetacea has very large rts, excluded them from analysis
median(rts_df$rts)
mean(rts_df$rts)

ggplot(rts_df, aes(rts)) +
  geom_histogram() +
  theme_classic()

# already removed species with incorrectly provided testes mass
#remove_species <- c("Cynomys_parvidens", "Funisciurus_congicus")





```





# Collinearity of variables
Prepare data
```{r}
## Collinearity of the explanatory variables ----
library(psych) #Calls: pairs.panels
library(car) #Calls: vif
library(plyr) #Calls: rbind.fill
#library(usdm)

# global, without relative testes size but with body mass and testes mass
# Z <- cbind(traits$BM_g,traits$testes_g, traits$LG, traits$LS , traits$LY )
# colnames(Z) <- c("bm","testes" ,"lg","ls","ly")
# pairs.panels(Z) # Generate scatterplots with Pearson correlations
# dfz = data.frame(Z) # Data Frame with predictor variables
# vif(dfz) # no variable with VIF > 3

# 
Z <- cbind(rts_df$rts, rts_df$longevity, rts_df$litter_size , rts_df$litter_year ) #, df$paternal_care , df$mating_system
colnames(Z) <- c("rts", "lg","ls","ly") #, "pc","ms"
pairs.panels(Z) # Generate scatterplots with Pearson correlations
dfz = data.frame(Z) # Data Frame with predictor variables
vif(dfz) # VIF > 3 : no variables


# global
Z <- cbind(rts_df$rts,rts_df$longevity, rts_df$litter_size , rts_df$litter_year,
           rts_df$paternal_care , rts_df$mating_system ) 
colnames(Z) <- c("rts", "lg","ls","ly","pc", "mc") #, "pc","ms"
pairs.panels(Z) # Generate scatterplots with Pearson correlations
dfz = data.frame(Z) # Data Frame with predictor variables
vif(dfz) # VIF > 3 : no variables

vif(lm(rts ~ litter_size + litter_year + longevity + mating_system + paternal_care, data=rts_df)) # no problems with collinearity



```
No problem with collinearity

# Plots

```{r}
speciesXorder <- rts_df %>%
  group_by(order) %>%
  tally()
```


```{r litter size}

range(rts_df$litter_size, na.rm = T)
fig.ls <- ggplot(rts_df, aes(x=litter_size, y=rts)) + 
  geom_point(aes(color=order)) +
  #geom_smooth( method = "lm") + 
  theme_classic() +
  scale_x_continuous(limits = c(0.5,10.0), 
                     breaks = 0:10) + 
  labs(x= "litter size", y = "relative testes size") +
  scale_colour_manual(values = colorXorder) 
  # geom_point(aes(color=order)) +
  # scale_color_brewer(palette = "Set1")

fig.ls
```

```{r litters/year}
range(rts_df$LY, na.rm = T)

fig.ly<- ggplot(rts_df, aes(x=litter_year, y=rts)) + 
  geom_point(aes(color=order)) +
  theme_classic() +
  scale_x_continuous(limits = c(0.0,8.0), 
                     breaks = 1:8) + 
  labs(x= "litters per year", y = "relative testes size") +
  scale_colour_manual(values = colorXorder) 
fig.ly
```


```{r longevity}

range(rts_df$LG, na.rm = T)
fig.lg<- ggplot(rts_df, aes(x=longevity, y=rts)) + 
  geom_point(aes(color=order)) +
  theme_classic() +
  # scale_x_continuous(limits = c(0.0,8.0), 
  #                    breaks = 0:8) + 
  labs(x= "longevity (years)", y = "relative testes size") +
  scale_colour_manual(values = colorXorder) 
fig.lg
```

```{r mating system}
unique(rts_df$mating_system)
check <- rts_df %>%
  filter(mating_system=="monogamous; polygynous")
#change to most complex one
rts_df$mating_system[rts_df$mating_system=="monogamous; polygynous"] <- "polygynous"

matingsys <- c("monogamous", "polygynous", "promiscuous")

# switch to raincloud plot
fig.ms <- rts_df %>%
  filter(mating_system %in% matingsys) %>%
  ggplot( aes(x=mating_system, y=rts)) + 
  geom_boxplot() + 
  #geom_jitter(aes(color=order))+
  theme_classic() +
  labs(x= "mating system", y = "relative testes size")+
  scale_colour_manual(values = colorXorder) 
fig.ms
```


```{r paternal care}
unique(rts_df$paternal_care)

fig.pc<- rts_df %>%
  filter(!is.na(paternal_care)) %>%
  ggplot( aes(x=paternal_care, y=rts)) + 
  geom_boxplot() +
  #geom_jitter(aes(color=order))+
  theme_classic() +
  labs(x= "paternal care", y = "relative testes size")+
  scale_colour_manual(values = colorXorder) 
fig.pc
```




```{r combine}
# plots
unique(df$order)

# scale_color_viridis(discrete = TRUE, option = "D")+
#   scale_fill_viridis(discrete = TRUE) 


library(ggpubr) 
plot.rts<-ggarrange(figA, figB, figC, figD,figE, labels = "auto")
plot.rts
ggexport(plot.rts, filename = "plot.rts.pdf")
```


# other rts analyses

```{r}
# log-transformed body and testes mass
rts_log <- pruned_df %>%
  mutate(
         exp.testes=0.035*log(BM_g)^0.72,
         rts=log(testes_g)/exp.testes) %>%
  dplyr::select(order, family, common_name, Binomial, rts, rel_testes_size, 
                exp.testes, testes_g, source_testes, everything())
```





```{r residuals of linear regression}


cor.test(log(rts_df$BM_g), log(rts_df$testes_g), method=c("pearson"))

lm.rts <- lm(log(testes_g) ~ log(BM_g) , data=rts_df)
summary(lm.rts)

rts_df$rts <- lm.rts$residuals 

ggplot(all_df, aes( y= log(testes_g),x =log(BM_g))) +
  geom_point() +
  theme_classic() +
  geom_smooth(method = "lm") +
  labs(x="log body mass (g)", y = "log testes mass (g)")
  



```

We found a significant linear correlation of log of testes mass against log body mass (R2 = 0.91, p < 0.001). 

-3.1 +- 0.69, F1, 235 = 1156, p < 0.001





