---
title: "2.Run models"
author: "Annemarie van der Marel"
date: "2022-12-15"
output: html_document
---
Here, we perform PGLMMS using the brms package as it can examine correlations between continuous and discrete traits, and can incorporate multiple measurements per species.

Analyses: 

monogamy with polygyny -> females do not mate multiply vs promiscuity -> females do mate multiply; models:
  - subset litter size of 1 
  - full dataset

Highlighting litter size of 1 with different mating system -> model subset by litter size of 1
assumption that litter of 1 --> there is no post-copulatory selection  -> confounding effect

Here we show the results of the species with litter size of 1

```{r}
sessionInfo()


#clear garbage; https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/gc
gc()



```



The version of RStan currently available on CRAN is not compatible with R4.2, and displays errors like the following:
"Error in compileCode(f, code, language = language, verbose = verbose) : 
Error in sink(type = "output") : invalid connection"

Instead, you will need to install the preview of rstan 2.26 using:
```{r}
install.packages("StanHeaders", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
install.packages("rstan", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
```
See https://github.com/stan-dev/rstan/wiki/Configuring-C---Toolchain-for-Windows

I changed back to R version 4.1.2 as I still received the error. 

# Load libraries
```{r setup, include=FALSE}
# data cleaning
library(tidyverse) 
library(job) # to run models in the background

# phylogenies
library(phytools)
library(ape)

# PGLMMs
library(brms)
library(rstan)
library(rlang)
library(Rcpp)
#library("loo")

#library(rethinking)

```
https://cran.r-project.org/web/packages/loo/vignettes/loo2-example.html#using-the-loo-package-for-model-checking-and-comparison 


Stan settings
```{r}
mycores <- parallel::detectCores() - 2

options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```


# Import data
```{r}

rts_df <- read.csv("../data/rts_life_mating_care.csv") %>%
  dplyr::select(-X)

species<- unique(rts_df$Binomial)
#write.csv(species, "species571.csv")


unique(rts_df$mating_system)
rts_df$mating_system[rts_df$mating_system=="monogamous; polygynous"] <- "polygynous"
unique(rts_df$order)

df.ls1 <- rts_df %>%
  select(order, family, common_name, Binomial, 
         rts, 
         litter_size, litter_year, longevity, 
         mating_system) %>%
  mutate(females_mate_multiply = if_else(mating_system=="monogamous", "no",
                                         if_else(mating_system=="polygynous", "no","yes")))



```


# select species trait data
We select species with data for all our predictor variables

```{r}
select_dfls1 <- na.omit(df.ls1) 
n_distinct(select_dfls1$Binomial)
n_distinct(select_dfls1$order)
```
## set up model species with trait data

Litter size of 1 for all species and only 1 species with paternal care, so exclude litter size and paternal care from model

standardize values
```{r}
dfls1 <- select_dfls1  %>% 
  filter(litter_size<=1) %>%  # should probably do <1, then 9 species extra
  mutate(across(c("litter_year", "longevity"), scale)) # standardize values
#st$Binomial
head(dfls1)

length(unique(dfls1$Binomial))

```

# distribution

```{r}
ggplot(select_dfls1, aes(rts)) +
  geom_histogram()

# shapiro test p >0.05, data normally distributed
#ggqqplot(rts_df, "rts")
shapiro.test(select_dfls1$rts)

range(select_dfls1$rts)


m <- lm(rts ~ litter_size, data = select_dfls1)
summary(m)
d <- data.frame(residuals = residuals(m),         # Residuals
                std_residuals = rstudent(m), # Studentized Residuals
                fitted = fitted(m),          # Fitted values
                cooks = cooks.distance(m))   # Cook's D
d <- mutate(d, observation = 1:nrow(d))           # Observation number

# histogram of residuals
ggplot(data = d, aes(x = std_residuals)) +
  geom_histogram(bins = 10)

# Check for Normality: QQ Normality plot of residuals
ggplot(data = d, aes(sample = std_residuals)) +
  stat_qq() +
  stat_qq_line()

# Check heteroscedasticity or constant variance
ggplot(d, aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0)

# Cook's D
ggplot(d, aes(x = observation, y = cooks)) +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dotted") +
  geom_hline(yintercept = 4/nrow(d), linetype = "dashed")
# outliers
d_outliers <- filter(d, cooks > 4/nrow(d))  # observation 86
#rts_df[59,]
rts_df[550,] # Sorex_araneus largest Cook's distance

#d_outliers # no outliers when log-transformed; 2,31,33,34


# distribution
fitdistrplus::plotdist(select_dfls1$rts)
fitdistrplus::descdist(select_dfls1$rts, discrete = FALSE)
fit.gamma <- fitdistrplus::fitdist(df$rts, 'gamma')
plot(fit.gamma)
fit.gamma$aic
fit.weibull <- fitdistrplus::fitdist(df$rts, "weibull")
plot(fit.weibull)
fit.weibull$aic
fit.norm <- fitdistrplus::fitdist(df$rts, "norm")
plot(fit.norm)
fit.norm$aic
fit.lnorm <- fitdistrplus::fitdist(df$rts, "lnorm")
plot(fit.lnorm)
fit.lnorm$aic

# (gamlss)
library(gamlss)
fit <- fitDist(select_dfls1$rts, k = 2, type = "realline", trace = FALSE, try.gamlss = TRUE)
summary(fit)
fit$fits  # best fit: c("SHASHo", "Sinh-Arcsinh")  
fit$failed



# brms
# fitting a brms model with a Gaussian likelihood
job::job({
model_normal <- brm(rts ~ 1, family = gaussian(), data = select_dfls1)

# fitting a brms model with a gamma
model_gamma <- brm(rts ~ 1, family = Gamma(), data =select_dfls1)

# fitting a brms model with a skew normal likelihood
model_skew <- brm(rts ~ 1, family = skew_normal(), data = select_dfls1)

# fitting a brms model with a log normal likelihood
model_lnorm <- brm(rts ~ 1, family = lognormal(), data = select_dfls1)

# fitting a brms model with a weibull likelihood
model_weib <- brm(rts ~ 1, family = weibull(), data = select_dfls1)
})

# posterior predictive checking
pp_check(model_normal, ndraws = 1e2)  
pp_check(model_gamma, ndraws = 1e2)
pp_check(model_skew, ndraws = 1e2)
pp_check(model_lnorm, ndraws = 1e2) # lognormal best fitting distribution
pp_check(model_weib, ndraws = 1e2) 
```


# Import phylogenetic tree

-	a mammalian phylogenetic tree
o	Upham et al. 2019 https://doi.org/10.1371/journal.pbio.3000494 
o	http://vertlife.org/phylosubsets/ website to select species or taxonomic groups or complete tree 
tree-pruner-812b7edc-05f3-4f07-9665-eb3a3cd861e0 for complete 571 species


```{r}
# Upham et al. 2019
#mammaltree <- read.nexus("./data/output.nex")
trees <- ape::read.nexus("../data/571mammals.nex")
```



## plot tree
```{r}

pdf(file = "./figures/sampletree.pdf", width = 9, height = 30)
plot(tree)
dev.off()
```

## check tree
```{r}
# species names in tree
sort(trees[[1]]$tip.label)

#mammaltree
#str(mammaltree)



TipsWant <- unique(select_df$Binomial) # 220 species that have all trait data

# Only include the species that are present in our dataset
tree.tr = drop.tip(tree, setdiff(tree$tip.label, TipsWant))
str(tree.tr)
plot.phylo(tree.tr)
tree.tr$tip.label
plot(tree.tr); axisPhylo()

# list of taxa with traits that are not in the phylogeny 
remove.table = setdiff(TipsWant, tree$tip.label)
remove.table # no species missing in phylogeny 

# remove from table all species that are not in phylogeny -> not necessary
# pruned_df = df[!(TipsWant %in% remove.table), ]
# #st<- st %>% filter(Binomial %in% TipsWant) %>% droplevels() # other way
# unique(pruned_df$Binomial)
# length(unique(pruned_df$Binomial))
```



# intercept only model

```{r}
# duplicate species list for mixed-effects model
dfls1$phylo<- dfls1$Binomial
```


filter species with trait data from phylogeny
```{r}
TipsWant.ls1 <- unique(dfls1$Binomial) # 71 species that have all trait data

pruned.tree.ls1 <- list() 
for(i in 1:100){
  pruned.tree.ls1[[i]]<- drop.tip(trees[[i]], setdiff(trees[[i]]$tip.label, TipsWant.ls1))
}
#sort(pruned.tree[[1]]$tip.label)
```

Covariance matrix
```{r}
# convert to covariance matrix (see https://cran.r-project.org/web/packages/brms/vignettes/brms_phylogenetics.html)
A.ls1<- list() 
for(i in 1:100){
  A.ls1[[i]]<- vcv.phylo(pruned.tree.ls1[[i]])
}
```




###Get prior 
https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations
http://www.stat.columbia.edu/~gelman/research/published/entropy-19-00555-v2.pdf
https://discourse.mc-stan.org/t/help-understanding-and-setting-informative-priors-in-brms/9574/12
sample_prior = "only" and then creating prior predictive plots via pp_check.
```{r}
prior.intercept.ls1<- brms::get_prior(rts ~ 1 +
            (1|gr(Binomial, cov = A.ls1[[1]])) +  (1|phylo), 
          data =  dfls1, #st_intercept, 
          data2 = list(A.ls1 = A.ls1[[1]]),
          family = lognormal())
prior.intercept.ls1

```

### model

model convergence diagnositcs:
"We may also investigate convergence numerically by means of the
scale reduction factor Rb (Gelman and Rubin 1992; Gelman et al. 2013; Vehtari et al. 2019), which should be close to one (i.e., R < b 1.05), 
and the effective sample size, which should be as large as possible but at least 400 to merely ensure reliable convergence diagnostics " Burkner (2021)
```{r}

#job::job({
 
m.intercept.ls1 <- brm(rts ~ 1 +
              (1|gr(Binomial, cov = A.ls1[[1]])) + (1|phylo), 
  data = dfls1, #st_intercept, 
   data2 = list(A.ls1 = A.ls1[[1]]),
  family = lognormal(), 
  prior = prior.intercept.ls1, #c(prior(student_t(3, 0, 2.5))) 
  chains = 4, cores = mycores, # sample_prior = TRUE / "only"
  iter = 5000, warmup = 3000,  # iter = 10000, warmup = 1500,
  control = list(adapt_delta = 0.99, max_treedepth = 15)
  #, save_pars = save_pars(all = TRUE)
  )
#})

# check summary, plot for diagnostics --> all good
#rhat(global)
#neff_ratio(global)
pairs(m.intercept.ls1)
summary(m.intercept.ls1, waic = TRUE)
#head(predict(global)) # predict responses based on the fitted model

# investigate model fit
loo(m.intercept.ls1) #, moment_match = TRUE
pp_check(m.intercept.ls1) 
plot(m.intercept.ls1, ask=FALSE)


# extract samples
ls1.intercept.samples<- posterior_samples(m.intercept.ls1)
ls1.intercept.draws <- as_draws(m.intercept.ls1)
ls1.intercept.loo <- loo(m.intercept.ls1)$estimate[3]

```



### loop for all 100 trees

```{r}


i=1
list <- c(1:100)

## loo
  ls1.intercept.looic  <- data.frame(runby=character(),
                                     looic=numeric())
## posterior samples
  samplesi <-  colnames(ls1.intercept.samples) 
  ls1.intercept.samples1 <- data.frame(matrix(nrow = 0, ncol = length(samplesi)))
  colnames(ls1.intercept.samples1) <- samplesi # assign column names


for(i in 1:100){
  runby <- list[i]
  print(runby)

  pruned.tree <- drop.tip(trees[[runby]], 
                          setdiff(trees[[runby]]$tip.label, TipsWant.ls1)) 
  A <- vcv.phylo(pruned.tree)

  
  m.intercepti <- brm(rts ~ 1 +
              (1|gr(Binomial, cov = A)) + (1|phylo), 
  data = dfls1, 
  family = lognormal(), 
  data2 = list(A = A),
  prior = prior.intercept.ls1, 
  chains = 4, cores = mycores, 
  iter = 5000, warmup = 3000,  
  control = list(adapt_delta = 0.99, max_treedepth = 15)
  )

  # model fit
  looic <- loo(m.intercepti)$estimate[3]
  pool <- cbind.data.frame(runby, 
                           looic)
  ls1.intercept.looic <-rbind(ls1.intercept.looic , pool )
  write.csv(ls1.intercept.looic,"../results/ls1.loo.intercept.looped.csv" )
  
  # extract samples
  ls1.intercept.samplesi<- posterior_samples(m.intercepti)
  ls1.post.intercept.looped<- rbind(ls1.intercept.samples1, ls1.intercept.samplesi)
  save(ls1.post.intercept.looped, file="../results/ls1.post.intercept.looped.robj")
  
  # extract samples
  # intercept.samplesi<- posterior_samples(m.intercepti)
  # intercept.post.looped<- rbind(intercept.samples, intercept.samplesi)
  # save( intercept.post.looped, file=" intercept.post.looped.robj")
  # 
  # intercept.drawsi <- as_draws(m.intercepti)
  # intercept.post.draw.looped<- rbind(intercept.draws, intercept.drawsi)
  # save(intercept.post.draw.looped, file="intercept.post.draw.looped.robj")
}

  


#save(interactioni, file="interactionmodel.looped.robj")
```



#  females multiple mate
## global model

### set up model
Litter size of 1 for all species and only 1 species with paternal care, so exclude litter size and paternal care from model



```{r}
# duplicate species list for mixed-effects model
dfls1$phylo<- dfls1$Binomial
```


filter species with trait data from phylogeny
```{r}
TipsWant.ls1 <- unique(dfls1$Binomial) # 88 species that have all trait data

pruned.tree <- list() 
for(i in 1:100){
  pruned.tree[[i]]<- drop.tip(trees[[i]], setdiff(trees[[i]]$tip.label, TipsWant.ls1))
}
#sort(pruned.tree[[1]]$tip.label)
```


Covariance matrix
```{r}
# convert to covariance matrix (see https://cran.r-project.org/web/packages/brms/vignettes/brms_phylogenetics.html)
A<- list() 
for(i in 1:100){
  A[[i]]<- vcv.phylo(pruned.tree[[i]])
}
```


Get prior 
https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations
http://www.stat.columbia.edu/~gelman/research/published/entropy-19-00555-v2.pdf
https://discourse.mc-stan.org/t/help-understanding-and-setting-informative-priors-in-brms/9574/12
sample_prior = "only" and then creating prior predictive plots via pp_check.
```{r}
prior.ls1.fmm <- brms::get_prior(rts ~  litter_year + females_mate_multiply +
                                  longevity + 
            (1|gr(Binomial, cov = A[[1]])) +  (1|phylo), 
          data = dfls1, 
          data2 = list(A = A[[1]]),
          family = lognormal())
prior.ls1.fmm

```


### model

model convergence diagnositcs:
"We may also investigate convergence numerically by means of the
scale reduction factor Rb (Gelman and Rubin 1992; Gelman et al. 2013; Vehtari et al. 2019), which should be close to one (i.e., R < b 1.05), 
and the effective sample size, which should be as large as possible but at least 400 to merely ensure reliable convergence diagnostics " Burkner (2021)
```{r}

#job::job({
global.fmm.ls1 <- brm(rts ~ litter_year + longevity +
                     females_mate_multiply + 
                     (1|gr(Binomial, cov = A[[1]])) + (1|phylo), 
  data = dfls1, 
  family = lognormal(), 
  data2 = list(A = A[[1]]),
  prior = prior.ls1.fmm, #c(prior(student_t(3, 0, 2.5))) 
  chains = 4, cores = mycores, # sample_prior = TRUE / "only"
  iter = 5000, warmup = 3000,  # iter = 10000, warmup = 1500,
  control = list(adapt_delta = 0.99, max_treedepth = 15)
  #, save_pars = save_pars(all = TRUE)
  )
#})

# check summary, plot for diagnostics --> all good
#rhat(global)
#neff_ratio(global)
pairs(global.fmm.ls1 )
summary(global.fmm.ls1 , waic = TRUE)
#head(predict(global)) # predict responses based on the fitted model

# investigate model fit
loo(global.fmm.ls1 ) #, moment_match = TRUE
pp_check(global.fmm.ls1 ) 
plot(global.fmm.ls1 , ask=FALSE)


# extract samples
fmm.globalls1.samples<- posterior_samples(global.fmm.ls1)
fmm.globalls1.draws <- as_draws(global.fmm.ls1)




```
### loop for all 100 trees

```{r}

#job::job({
i=1
list <- c(1:100)

# make empty data frame to put data in
## loo
fmm.ls1.looic  <- data.frame(runby=character(),
                            looic=numeric())

## posterior samples
  samplesls1 <-  colnames(fmm.globalls1.samples) 
  ls1.samples1 <- data.frame(matrix(nrow = 0, ncol = length(samplesls1)))
  colnames(ls1.samples1) <- samplesls1 # assign column names


for(i in 1:100){
  runby <- list[i]
  print(runby)

  pruned.tree <- drop.tip(trees[[runby]], setdiff(trees[[runby]]$tip.label, TipsWant.ls1))
  A <- vcv.phylo(pruned.tree)

  
  global.fmm.ls1i <- brm(rts ~ litter_year + longevity +
                     females_mate_multiply + 
                     (1|gr(Binomial, cov = A)) + (1|phylo), 
  data = dfls1, 
  family = lognormal(), 
  data2 = list(A = A),
  prior = prior.ls1.fmm, #c(prior(student_t(3, 0, 2.5))) 
  chains = 4, cores = mycores, # sample_prior = TRUE / "only"
  iter = 5000, warmup = 3000,  # iter = 10000, warmup = 1500,
  control = list(adapt_delta = 0.99, 
                 max_treedepth = 15)
  #, save_pars = save_pars(all = TRUE)
  )

  # model fit
  looic <- loo(global.fmm.ls1i)$estimate[3]
  pool <- cbind.data.frame(runby, 
                           looic)
  fmm.ls1.looic <-rbind(fmm.ls1.looic , pool )
  write.csv(fmm.ls1.looic ,"../results/ls1.loo.global.looped.csv" )
  
  # extract samples
  ls1.samplesi<- posterior_samples(global.fmm.ls1i)
  post.ls1.looped<- rbind(ls1.samples1 , ls1.samplesi)
  save(post.ls1.looped, file="../results/post.ls1.global.looped.robj")
  
}



```



## interaction model
We hypothesized that the relationship between testes size and female reproductive output would differ between mammal species exhibiting different mating systems.  Specifically, in species where females mate with multiple males per breeding attempt, we predicted larger testes size with smaller litter sizes and fewer total annual offspring, because smaller litter sizes represent more intense competition due to fewer potential offspring available to be sired. In species where females mate with a single male per breeding attempt, we predicted larger testes with larger litter sizes and greater numbers of total annual offspring, due to lesser competition with other males, but an increased need to fertilize eggs.

female reproductive investment = litter size * litters/year 


### set up model

Litter size of 1 for all species and only 1 species with paternal care, so exclude litter size and paternal care from model



```{r}
# duplicate species list for mixed-effects model
dfls1$phylo<- dfls1$Binomial
```


filter species with trait data from phylogeny
```{r}
TipsWant.ls1 <- unique(dfls1$Binomial) # 88 species that have all trait data

pruned.tree <- list() 
for(i in 1:100){
  pruned.tree[[i]]<- drop.tip(trees[[i]], setdiff(trees[[i]]$tip.label, TipsWant.ls1))
}
#sort(pruned.tree[[1]]$tip.label)
```


Covariance matrix
```{r}
# convert to covariance matrix (see https://cran.r-project.org/web/packages/brms/vignettes/brms_phylogenetics.html)
A<- list() 
for(i in 1:100){
  A[[i]]<- vcv.phylo(pruned.tree[[i]])
}
```



Get prior 
https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations
http://www.stat.columbia.edu/~gelman/research/published/entropy-19-00555-v2.pdf
https://discourse.mc-stan.org/t/help-understanding-and-setting-informative-priors-in-brms/9574/12
sample_prior = "only" and then creating prior predictive plots via pp_check.
```{r}
prior.interaction.ls1 <- brms::get_prior(rts ~  litter_year * females_mate_multiply +
                                  longevity + 
            (1|gr(Binomial, cov = A[[1]])) +  (1|phylo), 
          data = dfls1, 
          data2 = list(A = A[[1]]),
          family = lognormal())


```

### model

model convergence diagnositcs:
"We may also investigate convergence numerically by means of the
scale reduction factor Rb (Gelman and Rubin 1992; Gelman et al. 2013; Vehtari et al. 2019), which should be close to one (i.e., R < b 1.05), 
and the effective sample size, which should be as large as possible but at least 400 to merely ensure reliable convergence diagnostics " Burkner (2021)
```{r}

job::job({
m.int.ls1.fmm <- brm(rts ~  litter_year * females_mate_multiply +
              longevity + 
              (1|gr(Binomial, cov = A[[1]])) + (1|phylo), 
  data = dfls1, 
  family = lognormal(), 
  data2 = list(A = A[[1]]),
  prior = prior.interaction.ls1, #c(prior(student_t(3, 0, 2.5))) 
  chains = 4, cores = mycores, # sample_prior = TRUE / "only"
  iter = 5000, warmup = 3000,  # iter = 10000, warmup = 1500,
  control = list(adapt_delta = 0.99, max_treedepth = 15)
  #, save_pars = save_pars(all = TRUE)
  )
})

# check summary, plot for diagnostics --> all good
#rhat(global)
#neff_ratio(global)
pairs(m.int.ls1.fmm)
summary(m.int.ls1.fmm, waic = TRUE)
#head(predict(global)) # predict responses based on the fitted model

# investigate model fit
pp_check(m.int.ls1.fmm) 
plot(m.int.ls1.fmm, ask=FALSE)

loo(m.int.ls1.fmm) #, moment_match = TRUE
#loo.interaction <- loo(m.int.ls1.fmm)$estimate[3]

# extract samples
m.int.ls1.fmm.samples<- posterior_samples(m.int.ls1.fmm)
m.int.ls1.fmm.draws <- as_draws(m.int.ls1.fmm)
m.int.ls1.fmm.looic <- loo(m.int.ls1.fmm)$estimate[3]

```
### loop for all 100 trees

Have to rerun again as I made a mistake
```{r}

#job::job({
i=1
list <- c(1:100)

# make empty data frame to put data in
## loo
ls1.interaction.looic  <- data.frame(runby=character(),
                            looic=numeric())
## posterior samples
  int.samplesls1 <-  colnames(m.int.ls1.fmm.samples) 
  int.ls1.samples1 <- data.frame(matrix(nrow = 0, ncol = length(int.samplesls1)))
  colnames(int.ls1.samples1) <- int.samplesls1 # assign column names


for(i in 1:100){
  runby <- list[i]
  print(runby)

  pruned.tree <- drop.tip(trees[[runby]], setdiff(trees[[runby]]$tip.label, TipsWant.ls1))
  A <- vcv.phylo(pruned.tree)

  
  int.ls1i <- brm(rts ~ litter_year * females_mate_multiply + longevity + 
                     (1|gr(Binomial, cov = A)) + (1|phylo), 
  data = dfls1, 
  family = lognormal(), 
  data2 = list(A = A),
  prior = prior.interaction.ls1, #c(prior(student_t(3, 0, 2.5))) 
  chains = 4, cores = mycores, # sample_prior = TRUE / "only"
  iter = 5000, warmup = 3000,  # iter = 10000, warmup = 1500,
  control = list(adapt_delta = 0.99, 
                 max_treedepth = 15)
  #, save_pars = save_pars(all = TRUE)
  )

  # model fit
  looic <- loo(int.ls1i)$estimate[3]
  pool <- cbind.data.frame(runby, 
                           looic)
  ls1.interaction.looic  <-rbind(ls1.interaction.looic , pool )
  write.csv(ls1.interaction.looic  ,"../results/ls1.loo.interaction.looped.csv" )
  
  # extract samples
  int.ls1.samplesi<- posterior_samples(int.ls1i)
  post.int.ls1.looped<- rbind(int.ls1.samples1 , int.ls1.samplesi)
  save(post.int.ls1.looped, file="../results/ls1.post.interaction.looped.robj")
  
}



```




# mating system 

There is only one species with monogamous mating system and litter size of 1, so exclude these models from analysis
```{r}

dfsl1_ms <-select_dfls1 %>%
  filter(litter_size==1) %>%
  group_by(mating_system) %>%
  tally() 
```


## global model

### set up model

Litter size of 1 for all species and only 1 species with paternal care, so exclude litter size and paternal care from model

standardize values
```{r}
dfls1 <- select_dfls1  %>% 
  filter(litter_size==1) %>%
  mutate(across(c("litter_year", "longevity"), scale)) # standardize values
#st$Binomial
head(dfls1)

length(unique(dfls1$Binomial))



```

```{r}
# duplicate species list for mixed-effects model
dfls1$phylo<- dfls1$Binomial
```


filter species with trait data from phylogeny
```{r}
TipsWant.ls1 <- unique(dfls1$Binomial) # 71 species that have all trait data

pruned.tree.ls1 <- list() 
for(i in 1:100){
  pruned.tree.ls1[[i]]<- drop.tip(trees[[i]], setdiff(trees[[i]]$tip.label, TipsWant.ls1))
}
#sort(pruned.tree[[1]]$tip.label)
```

Covariance matrix
```{r}
# convert to covariance matrix (see https://cran.r-project.org/web/packages/brms/vignettes/brms_phylogenetics.html)
A.ls1<- list() 
for(i in 1:100){
  A.ls1[[i]]<- vcv.phylo(pruned.tree.ls1[[i]])
}
```



###Get prior 
https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations
http://www.stat.columbia.edu/~gelman/research/published/entropy-19-00555-v2.pdf
https://discourse.mc-stan.org/t/help-understanding-and-setting-informative-priors-in-brms/9574/12
sample_prior = "only" and then creating prior predictive plots via pp_check.
```{r}
prior.global.ls1 <- brms::get_prior(rts ~  litter_year +
                                      longevity + mating_system + 
            (1|gr(Binomial, cov = A.ls1[[1]])) +  (1|phylo), 
          data = dfls1, 
          data2 = list(A.ls1 = A.ls1[[1]]),
          family = lognormal())
prior.global.ls1



```

### model

model convergence diagnositcs:
"We may also investigate convergence numerically by means of the
scale reduction factor Rb (Gelman and Rubin 1992; Gelman et al. 2013; Vehtari et al. 2019), which should be close to one (i.e., R < b 1.05), 
and the effective sample size, which should be as large as possible but at least 400 to merely ensure reliable convergence diagnostics " Burkner (2021)
```{r}

job::job({
global.ls1 <- brm(rts ~  litter_year + longevity + mating_system + 
                     (1|gr(Binomial, cov = A.ls1[[1]])) + (1|phylo), 
  data = dfls1, 
  family = lognormal(), 
  data2 = list(A.ls1 = A.ls1[[1]]),
  prior = prior.global.ls1, #c(prior(student_t(3, 0, 2.5))) 
  chains = 4, cores = mycores, # sample_prior = TRUE / "only"
  iter = 5000, warmup = 3000,  # iter = 10000, warmup = 1500,
  control = list(adapt_delta = 0.99, max_treedepth = 15)
  #, save_pars = save_pars(all = TRUE)
  )
})

# check summary, plot for diagnostics --> all good
#rhat(global)
#neff_ratio(global)
pairs(global.ls1)
summary(global.ls1, waic = TRUE)
#head(predict(global)) # predict responses based on the fitted model

# investigate model fit
loo(global.ls1) #, moment_match = TRUE
pp_check(global.ls1) 
plot(global.ls1, ask=FALSE)


# extract samples
ls1.global.samples<- posterior_samples(global.ls1)
ls1.global.draws <- as_draws(global.ls1)




```



### loop for all 100 trees

```{r}

#job::job({
i=1
list <- c(1:100)

# make empty data frame to put data in
## loo
ls1.global.looic  <- data.frame(runby=character(),
                            looic=numeric())
## posterior samples
  samplesgls1 <-  colnames(ls1.global.samples) 
  ls1.global.samples1 <- data.frame(matrix(nrow = 0, ncol = length(samplesgls1)))
  colnames(ls1.global.samples1) <- samplesgls1 # assign column names


for(i in 1:100){
  runby <- list[i]
  print(runby)

  pruned.tree.ls1 <- drop.tip(trees[[runby]], 
                              setdiff(trees[[runby]]$tip.label, TipsWant.ls1))
  A.ls1 <- vcv.phylo(pruned.tree.ls1)

  
  ls1.globali <- brm(rts ~ litter_year + longevity +
                     mating_system + 
                     (1|gr(Binomial, cov = A.ls1)) + (1|phylo), 
  data = dfls1, 
  family = lognormal(), 
  data2 = list(A.ls1 = A.ls1),
  prior = prior.global.ls1, #c(prior(student_t(3, 0, 2.5))) 
  chains = 4, cores = mycores, # sample_prior = TRUE / "only"
  iter = 5000, warmup = 3000,  # iter = 10000, warmup = 1500,
  control = list(adapt_delta = 0.99, 
                 max_treedepth = 15)
  #, save_pars = save_pars(all = TRUE)
  )

  # model fit
  looic <- loo(ls1.globali)$estimate[3]
  pool <- cbind.data.frame(runby, 
                           looic)
  ls1.global.looic <-rbind(ls1.global.looic , pool )
  write.csv(ls1.global.looic,"../results/ls1.loo.global.looped.csv" )
  
  # extract samples
  ls1.global.samplesi<- posterior_samples(ls1.globali)
  ls1.post.global.looped<- rbind(ls1.global.samples1, ls1.global.samplesi)
  save(ls1.post.global.looped, file="../results/ls1.post.global.looped.robj")
  

}


  
nrow(ls1.post.global.looped)/4000 # check all 100 models added 

```


## interaction model
We hypothesized that the relationship between testes size and female reproductive output would differ between mammal species exhibiting different mating systems.  Specifically, in species where females mate with multiple males per breeding attempt, we predicted larger testes size with smaller litter sizes and fewer total annual offspring, because smaller litter sizes represent more intense competition due to fewer potential offspring available to be sired. In species where females mate with a single male per breeding attempt, we predicted larger testes with larger litter sizes and greater numbers of total annual offspring, due to lesser competition with other males, but an increased need to fertilize eggs.

female reproductive investment = litter size * litters/year 


### set up model
Litter size of 1 for all species and only 1 species with paternal care, so exclude litter size and paternal care from model

standardize values
```{r}
dfls1 <- select_dfls1  %>% 
  filter(litter_size==1) %>%
  mutate(across(c("litter_year", "longevity"), scale)) # standardize values
#st$Binomial
head(dfls1)

length(unique(dfls1$Binomial))



```

```{r}
# duplicate species list for mixed-effects model
dfls1$phylo<- dfls1$Binomial
```


filter species with trait data from phylogeny
```{r}
TipsWant.ls1 <- unique(dfls1$Binomial) # 71 species that have all trait data

pruned.tree.ls1 <- list() 
for(i in 1:100){
  pruned.tree.ls1[[i]]<- drop.tip(trees[[i]], setdiff(trees[[i]]$tip.label, TipsWant.ls1))
}
#sort(pruned.tree[[1]]$tip.label)
```

Covariance matrix
```{r}
# convert to covariance matrix (see https://cran.r-project.org/web/packages/brms/vignettes/brms_phylogenetics.html)
A.ls1<- list() 
for(i in 1:100){
  A.ls1[[i]]<- vcv.phylo(pruned.tree.ls1[[i]])
}
```


###Get prior 
https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations
http://www.stat.columbia.edu/~gelman/research/published/entropy-19-00555-v2.pdf
https://discourse.mc-stan.org/t/help-understanding-and-setting-informative-priors-in-brms/9574/12
sample_prior = "only" and then creating prior predictive plots via pp_check.
```{r}
prior.interaction.ls1 <- brms::get_prior(rts ~ litter_year * mating_system +
                                  longevity + 
            (1|gr(Binomial, cov = A.ls1[[1]])) +  (1|phylo), 
          data = dfls1, 
          data2 = list(A.ls1 = A.ls1[[1]]),
          family = lognormal())
prior.interaction.ls1

```

### model

model convergence diagnositcs:
"We may also investigate convergence numerically by means of the
scale reduction factor Rb (Gelman and Rubin 1992; Gelman et al. 2013; Vehtari et al. 2019), which should be close to one (i.e., R < b 1.05), 
and the effective sample size, which should be as large as possible but at least 400 to merely ensure reliable convergence diagnostics " Burkner (2021)
```{r}

job::job({
m.int.ls1 <- brm(rts ~  litter_year *mating_system +  longevity + 
              (1|gr(Binomial, cov = A.ls1[[1]])) + (1|phylo), 
  data = dfls1, 
  family = lognormal(), 
  data2 = list(A.ls1 = A.ls1[[1]]),
  prior = prior.interaction.ls1 , #c(prior(student_t(3, 0, 2.5))) 
  chains = 4, cores = mycores, # sample_prior = TRUE / "only"
  iter = 5000, warmup = 3000,  # iter = 10000, warmup = 1500,
  control = list(adapt_delta = 0.99, max_treedepth = 15)
  #, save_pars = save_pars(all = TRUE)
  )
})

 # check summary, plot for diagnostics --> all good
#rhat(global)
#neff_ratio(global)
pairs(m.int.ls1)
summary(m.int.ls1, waic = TRUE)
#head(predict(global)) # predict responses based on the fitted model

# investigate model fit
pp_check(m.int.ls1) 
plot(m.int.ls1, ask=FALSE)

loo(m.ls1)
loo(m.int.ls1) #, moment_match = TRUE
loo.interaction.m.int.ls1 <- loo(m.int.ls1)$estimate[3]

# extract samples
m.int.ls1.samples<- posterior_samples(m.int.ls1)
m.int.ls1.draws <- as_draws(m.int.ls1)
m.int.ls1.looic <- loo(m.int.ls1)$estimate[3]



```




### loop for all 100 trees

```{r}
start.time <- Sys.time()

#job::job({
i=1
list <- c(1:100)

# make empty data frame to put data in
## loo
inter.looic <- data.frame(runby=character(),
                          looic=numeric())
## posterior samples
  samples <-  colnames(inter.samples) 
  inter.samples1 <- data.frame(matrix(nrow = 0, ncol = length(samples)))
  colnames(inter.samples1) <- samples # assign column names

for(i in 1:100){
  runby <- list[i]
  print(runby)

  pruned.tree <- drop.tip(trees[[runby]], setdiff(trees[[runby]]$tip.label, TipsWant))
  A <- vcv.phylo(pruned.tree)

  
  interactioni <- brm(rts ~ litter_size * litter_year * mating_system +
              longevity + paternal_care +
              (1|gr(Binomial, cov = A)) + (1|phylo), 
  data = st, 
  family = lognormal(), 
  data2 = list(A = A),
  prior = prior.interaction, 
  chains = 4, cores = mycores, 
  iter = 5000, warmup = 3000,  
  control = list(adapt_delta = 0.99, max_treedepth = 15)
  )

  # model fit
  #inter.looi <- loo(interactioni)[["looic"]]
  looic <- loo(interactioni)$estimate[3]
  pool <- cbind.data.frame(runby, 
                           looic)
  inter.looic <-rbind(inter.looic, pool )
  write.csv(inter.looic,"../results/loo.interaction.looped.csv" )
  #save(inter.looi, file="loo.interaction.looped.robj")
  
  # extract samples
  inter.samplesi<- posterior_samples(interactioni)
  post.inter.looped<- rbind(inter.samples1, inter.samplesi)
  save(post.inter.looped, file="../results/post.interaction.looped_run1-100.robj")
  
  # # have to combine lists
  # inter.drawsi <- as_draws(interactioni)
  # post.inter.draw.looped<- rbind(sample.draws, inter.drawsi)
  # save(post.inter.draw.looped, file="post.inter.draw.looped_run1-100.robj")
}
#})
end.time <- Sys.time()

#Time to run:
end.time - start.time
  
nrow(post.inter.looped)/2000 # check all 100 models added 

#save(interactioni, file="interactionmodel.looped.robj")
```


