---
title: "visualize results"
author: "Annemarie van der Marel"
date: "2023-01-15"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

# Load libraries
```{r setup, include=FALSE}
# data cleaning
library(tidyverse) 

#plotting
library(viridis)
library(RColorBrewer)

```


# import data
 full dataset
```{r}
rts_df <- read.csv("../data/rts_life_mating_care.csv") %>%
  dplyr::select(-X)

unique(rts_df$mating_system)
rts_df$mating_system[rts_df$mating_system=="monogamous; polygynous"] <- "polygynous"
unique(rts_df$order)

df <- rts_df
select_df <- na.omit(df) 
```

litter size 1
```{r}
df.ls1 <- rts_df %>%
  select(order, family, common_name, Binomial, 
         rts, 
         litter_size, litter_year, longevity, 
         mating_system) %>%
  mutate(females_mate_multiply = if_else(mating_system=="monogamous", "no",
                                         if_else(mating_system=="polygynous", "no","yes")))


select_dfls1 <- na.omit(df.ls1) 
dfls1 <- select_dfls1  %>% 
  filter(litter_size==1) %>%
  mutate(across(c("litter_year", "longevity"), scale)) # standardize values
#st$Binomial
head(dfls1)
```



# Plots full dataset

```{r set colors by order}

n_distinct(select_df$order)

# 26 qualitatively different colors from Dave Keenan
colors26 <-  c("#00B7FF", "#004DFF", "#00FFFF", "#826400", "#580041", "#FF00FF", "#00FF00",
               "#C500FF", "#B4FFD7", "#FFCA00", "#969600", "#B4A2FF", "#C20078", "#0000C1",
               "#FF8B00", "#FFC8FF", "#666666", "#FF0000", "#CCCCCC", "#009E8F", "#D7A870", 
               "#8200FF", "#960000", "#BBFF00",  "#006F00", "#1f78b4", "#8E0152") # "#1f78b4", "#8E0152", "#FFFF00", are the extr 2 colors

set.seed(42)
colorXorder <-  sample(colors26, n_distinct(select_df$order))

```
## litters/year
```{r litters/year}
range(select_df$litter_year, na.rm = T)

fig.ly<- ggplot(select_df, aes(x=litter_year, y=rts)) + 
  geom_point(aes(color=order)) +
  theme_classic() +
  scale_x_continuous(limits = c(0.0,8.0), 
                     breaks = 1:8) + 
  labs(x= "litters per year", y = "relative testes size") +
  scale_colour_manual(values = colorXorder) +
  geom_abline(intercept=-1.08, slope= -0.93)
fig.ly

ggsave("../figures/rts~ly.pdf", width = 7, height = 5) 
```


## mating system
```{r mating system}
unique(select_df$mating_system)

# switch to raincloud plot
fig.ms <- 
  ggplot(select_df, aes(x=mating_system, y=rts)) + 
  geom_boxplot() + 
  #geom_jitter(aes(color=order))+
  theme_classic() +
  labs(x= "mating system", y = "relative testes size")+
  scale_colour_manual(values = colorXorder) 
fig.ms

ggsave( "../figures/rts~mating_system.pdf", width = 7, height = 5) 
```
## litters/year X mating system

```{r interaction}

unique(select_df$mating_system)

# 26 qualitatively different colors from Dave Keenan
colors_ms <-  c("#C20078", "#004DFF",   "#FFCA00") #,  "#006F00" "#1f78b4", "#8E0152", "#FFFF00", are the extr 2 colors



fig.ls.ms <- ggplot(select_df, aes(x=litter_year, y=rts)) + 
  geom_point(aes(color=mating_system)) +
  #geom_smooth( method = "lm") + 
  theme_classic() +
  scale_x_continuous(limits = c(0.5,10.0), 
                     breaks = 0:10) + 
  labs(x= "litters per year", y = "relative testes size") +
  scale_colour_manual(values = colors_ms) 
  # geom_point(aes(color=order))
fig.ls.ms

ggsave("../figures/rts~lyXms.pdf", width = 7, height = 5)


# just rodents
ggplot(filter(select_df, order=="Rodentia"), aes(x=litter_year, y=rts)) + 
  geom_point(aes(color=mating_system)) +
  #geom_smooth( method = "lm") + 
  theme_classic() +
  scale_x_continuous(limits = c(0.5,10.0), 
                     breaks = 0:10) + 
  labs(x= "litters per year", y = "relative testes size") +
  scale_colour_manual(values = colors_ms) 

```

# Plots litter size 1
Here we subsetted the dataset by species that have a litter size of 1
```{r females multiple mating}

fig.ls1.fmm <- ggplot(dfls1, aes(x=females_mate_multiply, y=rts)) +
  geom_boxplot() +
  theme_classic() +
  labs(x= "female multiple mating present", y = "relative testes size")
fig.ls1.fmm

ggsave( "../figures/ls1.rts~fmm.pdf", width = 7, height = 5) 

```


```{r litters/uear}


fig.ls1.ly <- ggplot(dfls1, aes(x=litter_year, y=rts)) +
  geom_point(aes(color=order)) +
  theme_classic() +
  labs(x= "litters per year", y = "relative testes size") +
  scale_colour_manual(values = colorXorder) 
  #scale_x_continuous(limits = c(0.0,8.0), breaks = 1:8)

fig.ls1.ly

ggsave( "../figures/ls1.rts~ly.png", width = 7, height = 5) 


```


# plots not included in manuscript
These factors had a negligible effect on relative testes size as the posterior distribution of the parameter crossed zero. 

## litter size

Potentially rerun models without the Tenrec_ecaudatus which have litter size of 16 
```{r litter size}



range(select_df$litter_size, na.rm = T)
fig.ls <- ggplot(select_df, aes(x=litter_size, y=rts)) + 
  geom_point(aes(color=order)) +
  #geom_smooth( method = "lm") + 
  theme_classic() +
  # scale_x_continuous(limits = c(0.5,10.0), 
  #                    breaks = 0:10) + 
  labs(x= "litter size", y = "relative testes size") +
  scale_colour_manual(values = colorXorder) 
  # geom_point(aes(color=order)) +
  # scale_color_brewer(palette = "Set1")

fig.ls

ggplot(filter(select_df, order=="Rodentia"), aes(x=litter_size, y=rts)) + 
  geom_point(aes(color=order)) +
  #geom_smooth( method = "lm") + 
  theme_classic() +
  # scale_x_continuous(limits = c(0.5,10.0), 
  #                    breaks = 0:10) + 
  labs(x= "litter size", y = "relative testes size") +
  scale_colour_manual(values = colorXorder) 
  # geom_point(aes(color=order)) +
  # scale_color_brewer(palette = "Set1")


ggsave("../figures/rts~ls.pdf", width = 7, height = 5)
```
## longevity
```{r longevity}

range(select_df$longevity, na.rm = T)
fig.lg<- ggplot(select_df, aes(x=longevity, y=rts)) + 
  geom_point(aes(color=order)) +
  theme_classic() +
  # scale_x_continuous(limits = c(0.0,8.0), 
  #                    breaks = 0:8) + 
  labs(x= "longevity (years)", y = "relative testes size") +
  scale_colour_manual(values = colorXorder) 
fig.lg
ggsave("../figures/rts~lg.pdf", width = 7, height = 5)
```
## parental care
```{r paternal care}
unique(select_df$paternal_care)

fig.pc<- 
  ggplot(select_df, aes(x=paternal_care, y=rts)) + 
  geom_boxplot() +
  #geom_jitter(aes(color=order))+
  theme_classic() +
  labs(x= "paternal care", y = "relative testes size")+
  scale_colour_manual(values = colorXorder) 
fig.pc
ggsave("../figures/rts~pc.pdf", width = 7, height = 5)
```
## litter size by litter/year

```{r}
fig.lsXly <- ggplot(select_df, aes(x = litter_size, y = rts, size = litter_year))+
  geom_point()+
  labs(x = 'litter size', y = 'relative testes size')+
  scale_size_continuous() +
  theme_classic()
  
  # geom_abline(aes(intercept=33.965, slope=-4.3985, linetype='-1SD litters per year'))+
  # geom_abline(aes(intercept=38.1208, slope=-5.854, linetype='Mean litters per year'))+
  # geom_abline(aes(intercept=42.2767, slope=-7.3095, linetype='+1SD litters per year'))+
  # scale_linetype_manual(values=c('dotted','dashed','solid'),
  #                       breaks=c('-1SD litters per year','Mean litters per year','+1SD litters per year'),
  #                       name='Simple\nSlope')+

fig.lsXly

ggsave('lsXly.png', width=6, height=6, unit='in', dpi=300)
```


## litter size by mating system

```{r litter size X mating system}

unique(select_df$mating_system)

# 26 qualitatively different colors from Dave Keenan
colors_ms <-  c("#C20078", "#004DFF",   "#FFCA00",  "#006F00") # "#1f78b4", "#8E0152", "#FFFF00", are the extr 2 colors



lsXms <- ggplot(select_df, aes(x=litter_size, y=rts)) + 
  geom_point(aes(color=mating_system)) +
  #geom_smooth( method = "lm") + 
  theme_classic() +
  scale_x_continuous(limits = c(0.5,10.0), 
                     breaks = 0:10) + 
  labs(x= "litter size", y = "relative testes size") +
  scale_colour_manual(values = colors_ms) 
  # geom_point(aes(color=order))
lsXms

ggsave('lsXms.png', width=6, height=6, unit='in', dpi=300)
```

