---
title: "mcmcGLMM"
author: "Annemarie van der Marel"
date: "10/08/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
Load libraries
```{r setup, include=FALSE}
library(TreeTools)
library(tidyverse)
library(ape)
library(MCMCglmm)
library(brms)
library(phytools)
sessionInfo()
```

# Import  data
```{r}
# species names (to include in tree)
species <- read.csv("./data/species.csv") %>%
  dplyr::select(-X)
head(species)


all_df <- read.csv("./data/testes_lifehistory_mating_care_all.csv") %>%
  dplyr::select(order, family, common_name, Binomial, BM_g, source_BM, testes_g, source_testes, 
                rel_testes_size, LG, LY, LS, source_lifehistory,
                paternal_care, source_paternal_care, mating_system, source_MS)

check_df <- all_df %>% 
  group_by(Binomial) %>%
  tally() %>%
  filter(n>1)

```

## number of species per trait
```{r}
n_testes <- all_df %>%
  filter(is.na(testes_g)) 
n_distinct(all_df$Binomial)

n_ls <- all_df %>%
  filter(is.na(LS)) 
n_distinct(n_ls$Binomial)
length(unique(check_df$Binomial))

n_ly <- all_df %>%
  filter(is.na(LY))
n_distinct(n_ly$Binomial)

n_lg <- all_df %>%
  filter(is.na(LG))
n_distinct(n_lg$Binomial)

n_ms <- all_df %>%
  filter(is.na(mating_system))
n_distinct(n_ms$Binomial)

n_pc <- all_df %>%
  filter( paternal_care=="")
n_distinct(n_pc$Binomial)

n_all <- na.omit(all_df) 
n_distinct(n_all)

```

in total, we have a dataset of 210 species with testes size or relative testes size information. 
Of these 210 species, we are missing litter size data for 12 species, litters per year data for 63 species, longevity data for 40 species, mating system data for 102 species, and paternal care data for 102 species. We have 12 species that are all squirrels with data for all variables. 




# import phylogenetic tree

-	a mammalian phylogenetic tree
o	Upham et al. 2019 https://doi.org/10.1371/journal.pbio.3000494 
o	http://vertlife.org/phylosubsets/ website to select species or taxonomic groups or complete tree 

or mammalian supertree from Rolland et al. 2014, see also Lukas & Clutton-brock 2020. 

```{r}
# Upham et al. 2019
#mammaltree <- read.nexus("./data/output.nex")
mammaltree <- ape::read.nexus("./data/output.nex")

# randomly choose 1 tree out of 1000
set.seed(42)
sample(1000,1)

tree <- mammaltree[[74]]
tree


# rolland et al 2014
Rollandtree <- phytools::read.newick("Rolland et al tree.txt")
#plot(Rollandtree)
str(Rollandtree)
tree<-Rollandtree
```


# check tree
```{r}
# species names in tree
sort(tree$tip.label)


pdf(file = "./figures/sampletree.pdf", width = 9, height = 30)
plot(tree)
dev.off()



mammaltree
str(mammaltree)

# only 204 species in tree instead 210
# species that have trait data
TipsWant <- rts_df$Binomial
# keep only the taxa we are interested on
tree.tr = drop.tip(tree, setdiff(tree$tip.label, TipsWant))
plot(tree.tr); axisPhylo()

# list of taxa with traits that are not in the phylogeny
remove.table = setdiff(TipsWant, tree$tip.label)
remove.table 
# remove from table all species that are not in phylogeny
pruned_df = all_df[!(TipsWant %in% remove.table), ]


```

Change:
Spermophilus_beecheyi -> Otospermophilus_beecheyi
Spermophilus_beldingi -> Urocitellus_beldingi



# relative testes size
Kenagy & Trombulak 1986
Testes mass = 0.035 x body mass^0.72

Y = -0.031X^0.77 for rodents 
rts of Capes is 2.83 (Anjos thesis)
Richardson's ground squirrels = 0.61 ± 0.06

```{r}
rts_df <- pruned_df %>% 
  filter(!is.na(testes_g)) %>%
  group_by(Binomial) %>%
  slice(1) 

cor.test(log(rts_df$BM_g), log(rts_df$testes_g), method=c("pearson"))

lm.rts <- lm(log(testes_g) ~ log(BM_g) , data=rts_df)
summary(lm.rts)

rts_df$rts <- lm.rts$residuals 

ggplot(all_df, aes( y= log(testes_g),x =log(BM_g))) +
  geom_point() +
  theme_classic() +
  geom_smooth(method = "lm") +
  labs(x="log body mass (g)", y = "log testes mass (g)")
  



```

We found a significant linear correlation of log of testes mass against log body mass (R2 = 0.91, p < 0.001). 

-3.1 +- 0.69, F1, 235 = 1156, p < 0.001
https://pavelmatos.files.wordpress.com/2019/10/comparativephylogenetics_r_tutorial-1.pdf
```{r}
rts_df[["BM_g"]] = log(rts_df[["BM_g"]]) # log convert body sizes
rts_df[["testes_g"]] = log(rts_df[["testes_g"]]) # log convert testes size
mass = rts_df[["BM_g"]] # data on body sizes
testes = rts_df[["testes_g"]] # data on testes sizes

# compare body mass and testes size -> not working! Error in pic(mass, tree) : length of phenotypic and of phylogenetic data do not match
pic.mass = pic(mass, tree)
pic.testes = pic(testes, tree)
plot(pic.testes, pic.mass)
fit.pic.MT = lm(pic.testes~pic.mass -1)
abline(fit.pic.MT, col = "red")
summary(fit.pic.MT)


```


# PGLMM 
brms can examine correlations between continuous and discrete traits, and can incorporate multiple measurements per species.


## Collinearity of variables
Prepare data
```{r}
## Collinearity of the explanatory variables ----
library(psych) #Calls: pairs.panels
library(car) #Calls: vif
library(plyr) #Calls: rbind.fill
library(usdm)

df<- all_df

# global, without relative testes size but with body mass and testes mass
Z <- cbind(df$BM_g,df$testes_g, df$LG, df$LS , df$LY )
colnames(Z) <- c("bm","testes" ,"lg","ls","ly")
pairs.panels(Z) # Generate scatterplots with Pearson correlations
dfz = data.frame(Z) # Data Frame with predictor variables
vif(dfz) # no variable with VIF > 3

# global
Z <- cbind(df$RTS, df$LG, df$LS , df$LY , df$paternal_care , df$mating_system)
colnames(Z) <- c("rts", "lg","ls","ly", "pc","ms")
pairs.panels(Z) # Generate scatterplots with Pearson correlations
dfz = data.frame(Z) # Data Frame with predictor variables
vif(dfz) # VIF > 3

# check with dependent and random factor
Z.all <- cbind(df$RTS, df$LG, df$LS , df$LY , df$Binomial )
colnames(Z.all) <- c("rts", "lg","ls","ly", "species")
pairs.panels(Z.all) # Generate scatterplots with Pearson correlations
dfz = data.frame(Z.all) 
vif(dfz) # VIF > 3


```



Plot
```{r}
# plots

figA<- ggplot(df, aes(x=LS, y=RTS)) + 
  geom_point() +
  theme_classic() +
  labs(x= "litter size", y = "relative testes size") #+
  geom_point(aes(group = Binomial)) +
  geom_line(aes(group = Binomial),
            alpha = 0.5, colour = "darkgrey")
  
figB<- ggplot(df, aes(x=LY, y=RTS)) + 
  geom_point() +
  theme_classic() +
  labs(x= "litters per year", y = "relative testes size") 

figC<- ggplot(df, aes(x=LG, y=RTS)) + 
  geom_point() +
  theme_classic() +
  labs(x= "longevity (years)", y = "relative testes size") 

figD<- ggplot(df, aes(x=mating_system, y=RTS)) + 
  geom_boxplot() +
  theme_classic() +
  labs(x= "mating system", y = "relative testes size")

figE<- ggplot(df, aes(x=paternal_care, y=RTS)) + 
  geom_boxplot() +
  theme_classic() +
  labs(x= "paternal care", y = "relative testes size")

library(ggpubr) 
plot.rts<-ggarrange(figA, figB, figC, figD,figE, labels = "auto")
ggexport(plot.rts, filename = "plot.rts.pdf")
```


# Phylogenetic mixed model

standardize values
```{r}

df_scaled<-df %>% mutate_if(is.numeric, scale) #scale(x, center = TRUE, scale = TRUE)
head(st)

```


Phylogenetic tree doesn't include "Callosciurus_erythreus" & "Sciurus_aberti",
therefore exclude resulting in 11 species instead of 13 for our analysis. 

```{r}
st$Binomial<-as.factor(st$Binomial)
unique(st$Binomial)
st<- st %>% filter(!Binomial %in% c("Callosciurus_erythreus", "Sciurus_aberti")) %>% droplevels()
unique(st$Binomial)


```


Only include the species that are present in our dataset.
```{r}
pruned.tree<-drop.tip(mammaltree, setdiff(mammaltree$tip.label, species)) 
str(pruned.tree)
plot.phylo(pruned.tree)
pruned.tree$tip.label
```


Reorder data according to tips
```{r}
species<-pruned.tree$tip.label
st_tree_select<-dplyr::select(st, Phylo, Binomial,spread, Founders, WA, LY, AFB) 
st_tree_select$Binomial

traits<-  st_tree_select %>% arrange(factor(Binomial, levels = species))
traits$Binomial
```

# repeated measures brms

```{r}
A <- ape::vcv.phylo(pruned.tree) # covariance matrix
A <- ape::vcv.phylo(tree.tr) # covariance matrix
```


We only have repeated measures of testes mass and body mass of 23 species, so we will get rts data for 23 species. Are we taking the mean of 

```{r}
# get mean of repeated measures 
st_select_mean<-df
st_select_mean$Binomial
st_select_mean$mean_RTS <- 
  with(df, sapply(split(RTS, Binomial), mean)[Binomial])

st_select_mean$mean_BM <- 
  with(df, sapply(split(BM_g, Binomial), mean)[Binomial])
st_select_mean$mean_TM <- 
  with(df, sapply(split(testes_g, Binomial), mean)[Binomial])

head(st_select_mean)
```

get prior
```{r}
brms::get_prior(mean_TM ~ mean_BM  + 
            (1|gr(Binomial, cov = A))+ + (1|Phylo), 
          data = st_select_mean, 
          family = gaussian())


brms::get_prior(RTS ~ LS  + 
            (1|gr(Binomial, cov = A))+ + (1|Phylo), 
          data = st_select_mean, 
          family = gaussian())
get_prior(spread ~ mean_founders + mean_WA + mean_LY + mean_AFB + 
            (1|gr(Binomial, cov = A))+ + (1|Phylo), 
          data = st_select_mean, 
          family = bernoulli())
```

# model

```{r}
## model
model_repeat <- brm(
  spread ~ mean_founders + mean_WA + mean_LY + mean_AFB + 
    (1|gr(Binomial, cov = A))+ (1|Phylo), 
  data = st_select_mean, 
  family = bernoulli(), 
  data2 = list(A = A),
  prior = c(
    prior(student_t(3, 0, 2.5))), 
  sample_prior = TRUE, chains = 3, cores = 2, # results
  iter = 10000, warmup = 1500,
  control = list(adapt_delta = 0.99)
)

# checkout results: parameter settings
#sample_prior = TRUE, chains = 2, cores = 2, # checkout results
#  iter = 4000, warmup = 1000,
#  control = list(adapt_delta = 0.99)


pairs(model_repeat)
summary(model_repeat, waic = TRUE)
head(predict(model_repeat)) # predict responses based on the fitted model
plot(conditional_effects(model_repeat), ask = FALSE) # plot conditional effects for each predictor

# investigate model fit
loo(model_repeat) 
pp_check(model_repeat) 


# get variables
#get_variables(model_repeat)
```



## add variability of the independent variables (predictor) within species
```{r}
# as only founders change within species, we only have to add founders
st_select_mean$within_founders <- st_select_mean$Founders - st_select_mean$mean_founders

model_repeat_within <- brm(
  spread ~ within_founders + mean_founders + mean_WA +  mean_LY + mean_AFB + 
    (1|gr(Binomial, cov = A))+ (1|Phylo), 
  data = st_select_mean, 
  family = bernoulli(), 
  data2 = list(A = A),
  #prior = c(
  #  prior(student_t(3, 0, 2.5))),
  sample_prior = TRUE, chains = 3, cores = 2, #  results
  iter = 10000, warmup = 1500,
  control = list(adapt_delta = 0.999, max_treedepth = 15),
  save_all_pars = TRUE
) 

 sample_prior = TRUE, chains = 2, cores = 2, # checkout results
  iter = 4000, warmup = 1000,
  control = list(adapt_delta = 0.99, max_treedepth = 15),
  save_all_pars = TRUE

# optional
  sample_prior = TRUE, chains = 3, cores = 2, 
  iter = 1500000, warmup = 500000, thin = 10,
  control = list(adapt_delta = 0.999, max_treedepth = 15)
)

summary(model_repeat_within)
plot(model_repeat_within)
pairs(model_repeat_within)
summary(model_repeat_within, waic = TRUE)
head(predict(model_repeat_within)) # predict responses based on the fitted model
plot(conditional_effects(model_repeat_within), ask = FALSE) # plot conditional effects for each predictor

# investigate model fit
loo(model_repeat_within) 
pp_check(model_repeat_within) 
```


# model selection
```{r}
loo(model_repeat_within, model_repeat) # model repeat lowest looic
loo_compare(model_repeat_within, model_repeat, criterion = "waic")
waic(model_repeat_within, model_repeat) # loo better for the models
```

# plot posterior distribution

plotting posterior distributions
see https://mran.microsoft.com/snapshot/2020-02-15/web/packages/ggdistribute/vignettes/geom_posterior.html 
https://cran.r-project.org/web/packages/bayesplot/vignettes/plotting-mcmc-draws.html
```{r}
plot.model<- plot(model_repeat)
ggexport(plot.model, filename = "plot.phylogenetic.model.pdf")
# geom_vline(x=0, linetype = "dashed")
     #pars = "^b_", combo = c("dens", "trace")

# plotting posterior distributions
library(ggdistribute)
install.packages("ggdistribute")

# option bayesplot
library("bayesplot")

posterior <- as.array(model_repeat)
dim(posterior)
dimnames(posterior)

color_scheme_set("red")
mcmc_intervals(posterior)

color_scheme_set("teal")
post.plot<-mcmc_intervals(posterior, pars = c("b_Intercept", "b_mean_founders", "b_mean_WA", "b_mean_LY", "b_mean_AFB"), point_est="none")
post.plot

#The default is to show 50% intervals (the thick segments) and 90% intervals (the thinner outer lines). These defaults can be changed using the prob and prob_outer arguments, respectively. The points in the above plot are posterior medians. 
mcmc_areas(
  posterior, 
  pars = c("b_Intercept", "b_mean_founders", "b_mean_WA", "b_mean_LY", "b_mean_AFB"),
  prob = 0.8, # 0.8 = 80% intervals
  prob_outer = 0.99, # 99%
  point_est = "none"
)

mcmc_areas_ridges(
  posterior, 
  pars = c("b_Intercept", "b_mean_founders", "b_mean_WA", "b_mean_LY", "b_mean_AFB"),
  prob = 0.5, # 0.8 = 80% intervals
  prob_outer = 0.95, # 99%
  point_est = "median"
)

#The mcmc_hist function plots marginal posterior distributions (combining all chains) via histograms
color_scheme_set("green")
mcmc_hist(posterior, pars = c("b_Intercept", "b_mean_founders", "b_mean_WA", "b_mean_LY", "b_mean_AFB"))

color_scheme_set("brightblue")
mcmc_hist_by_chain(posterior, pars = c("b_Intercept", "b_mean_founders", "b_mean_WA", "b_mean_LY", "b_mean_AFB"))

#  plots kernel density estimates 
color_scheme_set("teal")
mcmc_dens(posterior, pars = c("b_Intercept", "b_mean_founders", "b_mean_WA", "b_mean_LY", "b_mean_AFB"))

mcmc_dens_overlay(posterior, pars = c("b_Intercept", "b_mean_founders", "b_mean_WA", "b_mean_LY", "b_mean_AFB"))

# violin density plots
color_scheme_set("teal")
mcmc_violin(posterior, pars = c("b_Intercept", "b_mean_founders", "b_mean_WA", "b_mean_LY", "b_mean_AFB"), probs = c(0.1, 0.5, 0.9))

# 
color_scheme_set("pink")
mcmc_pairs(posterior, pars = c("b_Intercept", "b_mean_founders", "b_mean_WA", "b_mean_LY", "b_mean_AFB"),
           off_diag_args = list(size = 1.5))


## option 
library(ggmcmc)

S<- ggs(model_repeat)
S
str(S)
ggmcmc(S)
ggsggs_caterpillar(model_repeat)
```



# explanation model output
Solutions for warning messages
    - control = list(max_treedepth = 15) # <x> a positive integer  > default of 10
    - control = list(adapt_delta = 0.99) # <x> > default of 0.8 (require 0<delta<1)
    
Rhat: If Rhat is considerably greater than 1 (i.e., > 1.1), the chains have not yet 
converged and it is necessary to run more iterations and/or set stronger priors.

(1|gr(Binomial, cov = A)) : accounts for phylogenetic relationship between species
(1|Phylo) : accounts for any specific effect that would be independent of the phylogenetic relationship between species (e.g., environmental or niche effects).

There is a significant effect when credibility interval does not contain 0. 

Investigate model fit
 - loo(model_repeat); leave-one-out cross-validation for direct numerical comparison of model fit, smaller values of loo indicate better fit
 - pp_check(model_repeat); posterior predictive check.